---
title: "MeMoSLAP_RScript_CheckroutinesXNATData"
author: "Sven Paßmann"
date: "2024-08-12"
output: html_document
---

Loading necessary packages

```{r}
# install packages first if neceassary
# install.packages("readxl")
# install.packages("dplyr")

library(readxl)
library(dplyr)
```

remove previous calculations and figures

```{r}
rm(list = ls())
graphics.off()
```

------------------------------------------------------------------------

set the working directory path

```{r}
setwd("Y:/01_Studien/00_FOR5429_MeMoSLAP/RU_IT/Server/02_Checking_Routines")
```

read in the data

```{r}
Check_quest <- read_excel ("Y:/01_Studien/00_FOR5429_MeMoSLAP/RU_Experiments_Participants/11_PreliminaryAnalysis/Data/MeMoSLAP_Data_Neuropsych_Questionnaires.xlsx")

# rename column(s)
names(Check_quest)[names(Check_quest) == "sub-ID"] <- "participant_id"
names(Check_quest)[names(Check_quest) == "Project"] <- "project"

# reduce the dataframe
Check_quest <- Check_quest [, c(1:69)]
Check_quest <- subset (Check_quest, in_exclude == "included")

# convert in number
Check_quest$MWT <- as.numeric(Check_quest$MWT)
```

------------------------------------------------------------------------

Checking data gaps for ages and gender

*creates a file which contains the respective subject IDs where data is missing for the respective item*

```{r}
Check_age_gender <- Check_quest [, c(1, 6, 7)]

# creating empty file for output
file_name <- "Output/MeMoSLAP_Report_missdata_AgeGender_P137.txt"
cat("", file = file_name, append = FALSE)  # Create or overwrite the file


# loop through columns, write subjects with missing values separately for each
# column in txt-file
for (col in names(Check_age_gender)[-1]) {  # Exclude the 'subject' column
  missing_subjects <- Check_age_gender[is.na(Check_age_gender[[col]]), "participant_id"]
  
  if (length(missing_subjects) > 0) {
    cat("Subjects with missing data in", col, ":\n", file = file_name, append = TRUE)
    
    # Print each missing subject on a new line
    for (subject in missing_subjects) {
      cat(subject, "\n", file = file_name, append = TRUE)
    }
    
  } else {
    cat("No missing data in", col, ".\n", file = file_name, append = TRUE)
  }
  
  cat("\n", file = file_name, append = TRUE)  # Add a blank line between columns
}
```

------------------------------------------------------------------------

Checking data gaps for neuropsych file

```{r}
# checking neuropsych
Check_neuropsych <- Check_quest [, c(1, 33:69)]
```

1.  checking for valid scores

*creates a file which contains a list displaying valid scores as TRUE for the respective item*

```{r}
# Handedness
Check_neuropsych$Hand_check <- Check_neuropsych$handedness == 2


# BDI 
Check_neuropsych <- Check_neuropsych %>% mutate(BDI_check = between(1,
0, 20))


# VLMT 
Check_neuropsych <- Check_neuropsych %>% mutate(VLMT_DG1_check =
between(7, 0, 15))

Check_neuropsych <- Check_neuropsych %>% mutate(VLMT_DG2_check = between(8, 0,
15))

Check_neuropsych <- Check_neuropsych %>% mutate(VLMT_DG3_check = between(9, 0,
15))

Check_neuropsych <- Check_neuropsych %>% mutate(VLMT_DG4_check = between(10,
0, 15))

Check_neuropsych <- Check_neuropsych %>% mutate(VLMT_DG5_check = between(11,
0, 15))

Check_neuropsych <- Check_neuropsych %>% mutate(VLMT_inter_check = between(12,
0, 15))

Check_neuropsych <- Check_neuropsych %>% mutate(VLMT_DG6_check = between(13,
0, 15))

Check_neuropsych <- Check_neuropsych %>% mutate(VLMT_DG7_check = between(14,
0, 15))

Check_neuropsych <- Check_neuropsych %>% mutate(VLMT_DG8_c_check = between(15,
0, 15))

Check_neuropsych$VLMT_DG8_fptotal_check <- Check_neuropsych$VLMT_DG8_fptotal <= 15

Check_neuropsych$VLMT_DG8_fplistB_check <- Check_neuropsych$VLMT_DG8_fptotal <= 15

Check_neuropsych$VLMT_DG8_fpsemant_check <- Check_neuropsych$VLMT_DG8_fpsemant <= 15

Check_neuropsych$VLMT_DG8_fpphonem_check <- Check_neuropsych$VLMT_DG8_fpphonem <= 15


# Rey Figure 
Check_neuropsych <- Check_neuropsych %>% mutate(Rey_copy_check =
between(20, 0, 36)) 

Check_neuropsych <- Check_neuropsych %>%
mutate(Rey_immediate_check = between(21, 0, 36)) 

Check_neuropsych <-
Check_neuropsych %>% mutate(Rey_delayed_check = between(22, 0, 36))


# Digit span 
Check_neuropsych$Digit_forward_check <- Check_neuropsych$Digit_forward <= 12
 #Check_neuropsych %>%  mutate(Digit_forward_check = between(23, 0, 12)) 

Check_neuropsych$Digit_back_check <-  Check_neuropsych$Digit_back <= 12
 #Check_neuropsych %>% mutate(Digit_back_check = between(24, 0, 12))


# MWT 
Check_neuropsych <- Check_neuropsych %>% mutate(MWT_check = between(36,
0, 37))


# Corsi Block Tapping Task 
Check_neuropsych$Corsi_forward_check <-  Check_neuropsych$Corsi_forward <= 7
 # Check_neuropsych %>% mutate(Corsi_forward_check = between(37, 0, 7)) 

Check_neuropsych$Corsi_back_check <-  Check_neuropsych$Corsi_back <= 8
 # Check_neuropsych %>% mutate(Corsi_back_check = between(38, 0, 8))


# creating excel file with TRUE/FALSE scores for valid entries
Check_valid <- Check_neuropsych [, c(1, 39:61)]

write.table(Check_valid, file = "Output/MeMoSLAP_Report_CheckValidNeuropsych_P137.text", sep = "\t", row.names = FALSE, quote = FALSE)
```

2.  checking for gaps

*creates a file which contains the respective subject IDs where data is missing for the respective item*

```{r}
# reduce dataframe to columns of interest
Check_neuropsych <- Check_neuropsych[c(1:38)]


# creating empty file for output
file_name <- "Output/MeMoSLAP_Report_missdata_Neuropsych_P137.txt"
cat("", file = file_name, append = FALSE)  # Create or overwrite the file


# loop through columns, write subjects with missing values separately for each
# column in txt-file
for (col in names(Check_neuropsych)[-1]) {  # Exclude the 'subject' column
  missing_subjects <- Check_neuropsych[is.na(Check_neuropsych[[col]]), "participant_id"]
  
  if (length(missing_subjects) > 0) {
    cat("Subjects with missing data in", col, ":\n", file = file_name, append = TRUE)
    
    # Print each missing subject on a new line
    for (subject in missing_subjects) {
      cat(subject, "\n", file = file_name, append = TRUE)
    }
    
    } else {
    cat("No missing data in", col, ".\n", file = file_name, append = TRUE)
  }
  
  cat("\n", file = file_name, append = TRUE)  # Add a blank line between columns
}
```
